---
title: "final"
output: html_document
---

```{r}
library(dplyr)
library(ggplot2)
library(lme4)
```


```{r}
nba <- read.csv("nba_shots_combined.csv", stringsAsFactors = TRUE)
head(nba)
```

## Transformations

```{r}
nba <- nba %>%
  mutate(
    GAME_DATE = as.Date(as.character(GAME_DATE), format = "%Y%m%d"),
    HEIGHT = as.integer(str_split_fixed(HEIGHT, "-", 2)[,1]) * 12 +
                     as.integer(str_split_fixed(HEIGHT, "-", 2)[,2]),
    
  ) %>%
  select(
    -c(PLAYER_NAME_y, SEASON, TEAM_TEAM_NAME, GRID_TYPE)
  ) 
```

Lets get rid of players with less than 20 shots
```{r}
nba <- nba %>%
  group_by(PLAYER_ID) %>%
  filter(n() >= 20) %>%
  ungroup()
```


Add home team column:
```{r}

```



## Base EDA

Does not include all the games - API limited how much I could take at once - about half of the regular season games, all the  playoff games
```{r}
nba %>%
  summarize(
    num_players = n_distinct(PLAYER_ID),
    num_teams = n_distinct(TEAM_ID),
    num_games = n_distinct(GAME_ID),
    num_position = n_distinct(POSITION),
  )
```

Missing games:
```{r}
nba %>%
  filter(GAME_TYPE == "Regular Season") %>%
  group_by(GAME_DATE) %>%
  summarize(n_games = n_distinct(GAME_ID)) %>%
  arrange(GAME_DATE)
```

Mainly have start of the season games - cut off started in mid january.

How many shots per player? by team?
```{r}
nba %>%
  group_by(TEAM_ABBREVIATION, POSITION) %>%
  summarize(
    games_played = n_distinct(GAME_ID),
    total_shots = n(),
    avg_shots= total_shots / games_played,
    fg_pct = mean(SHOT_MADE_FLAG),
    avg_distance = mean(SHOT_DISTANCE),
    three_pt_rate = mean(SHOT_TYPE == "3PT Field Goal"),
    .groups = "drop"
  ) %>%
  arrange(desc(TEAM_ABBREVIATION))
```

```{r}
# most shots is SGA
nba %>%
  group_by(PLAYER_ID) %>%
  summarize(
    total_shots = n(),
    shots_made = sum(SHOT_MADE_FLAG),
    pct = shots_made / total_shots
  ) %>%
  summarize(
    avg_shots = mean(total_shots),
    avg_made = mean(shots_made),
    avg_pct = avg_made/avg_shots
  )
```



```{r}
# most shots is SGA
nba %>%
  group_by(PLAYER_ID) %>%
  summarize(
    total_shots = n(),
    shots_made = sum(SHOT_MADE_FLAG),
    pct = shots_made / total_shots
  ) %>%
  ggplot(aes(x=total_shots, y=shots_made)) +
  geom_point()
```

Looking at shot percentage against distance (binned) - looks decently linear
```{r}
nba %>%
  mutate(bin = ntile(SHOT_DISTANCE, 500)) %>%
  group_by(bin) %>%
  summarize(
    mean_distance = mean(SHOT_DISTANCE),
    prob = mean(SHOT_MADE_FLAG),
    odds = prob/(1-prob),
    logodds = log(odds)
  ) %>%
  ggplot(aes(x=mean_distance, y=logodds)) +
  geom_point()
```

Shot makes (filtered by jumpshot)
```{r}
nba %>%
  filter(ACTION_TYPE == "Jump Shot" | ACTION_TYPE == "Pullup Jump shot") %>%
  mutate(bin = ntile(SHOT_DISTANCE, 500)) %>%
  group_by(bin) %>%
  summarize(
    mean_distance = mean(SHOT_DISTANCE),
    prob = mean(SHOT_MADE_FLAG),
    odds = prob/(1-prob),
    logodds = log(odds)
  ) %>%
  ggplot(aes(x=mean_distance, y=logodds)) +
  geom_point()
```


```{r}
nba %>%
  filter(ACTION_TYPE == "Jump Shot" | ACTION_TYPE == "Pullup Jump shot") %>%
  group_by(GAME_ID) %>%
  summarize(
    mean_distance = mean(SHOT_DISTANCE),
    prob = mean(SHOT_MADE_FLAG),
    odds = prob/(1-prob),
    logodds = log(odds)
  ) %>%
  ggplot(aes(x=mean_distance, y=logodds)) +
  geom_point()
```

```{r}
nba %>%
  filter(ACTION_TYPE == "Jump Shot" | ACTION_TYPE == "Pullup Jump shot") %>%
  group_by(GAME_ID) %>%
  summarize(
    mean_distance = mean(SHOT_DISTANCE),
    prob = mean(SHOT_MADE_FLAG),
    odds = prob/(1-prob),
    logodds = log(odds)
  ) %>%
  ggplot(aes(x=logodds)) +
  geom_histogram()
```





Looking at ACTION_TYPE shots
```{r}
table(nba$ACTION_TYPE) 
```

```{r}
nba <- nba %>%
  mutate(SHOT_CATEGORY = case_when(
    ACTION_TYPE %in% c(
      "Alley Oop Layup shot", "Cutting Finger Roll Layup Shot", "Cutting Layup Shot",
      "Driving Finger Roll Layup Shot", "Driving Layup Shot", "Driving Reverse Layup Shot",
      "Finger Roll Layup Shot", "Layup Shot", "Putback Layup Shot", "Reverse Layup Shot",
      "Running Alley Oop Layup Shot", "Running Finger Roll Layup Shot", "Running Layup Shot",
      "Running Reverse Layup Shot", "Tip Layup Shot"
    ) ~ "Layup",
    
    ACTION_TYPE %in% c(
      "Alley Oop Dunk Shot", "Cutting Dunk Shot", "Driving Dunk Shot", "Driving Reverse Dunk Shot",
      "Dunk Shot", "Putback Dunk Shot", "Reverse Dunk Shot", "Running Alley Oop Dunk Shot",
      "Running Dunk Shot", "Running Reverse Dunk Shot", "Tip Dunk Shot"
    ) ~ "Dunk",
    
    ACTION_TYPE %in% c(
      "Driving Bank Hook Shot", "Driving Hook Shot", "Hook Bank Shot", "Hook Shot",
      "Turnaround Bank Hook Shot", "Turnaround Hook Shot"
    ) ~ "Hook Shot",
    
    ACTION_TYPE %in% c(
      "Fadeaway Bank shot", "Fadeaway Jump Shot", "Turnaround Fadeaway Bank Jump Shot",
      "Turnaround Fadeaway shot"
    ) ~ "Fadeaway",
    
    ACTION_TYPE %in% c(
      "Driving Floating Bank Jump Shot", "Driving Floating Jump Shot", "Floating Jump shot",
      "Jump Bank Shot", "Jump Shot", "Pullup Jump shot", "Running Jump Shot",
      "Running Pull-Up Jump Shot", "Step Back Bank Jump Shot", "Step Back Jump shot",
      "Turnaround Bank shot", "Turnaround Jump Shot"
    ) ~ "Jump Shot",
    
    TRUE ~ "Other"
  ))

table(nba$SHOT_CATEGORY)
```



```{r}
ggplot(nba, aes(x=factor(SHOT_ATTEMPTED_FLAG), fill = factor(SHOT_MADE_FLAG))) +
  facet_wrap(~factor(SHOT_CATEGORY)) +
  geom_bar(position = "fill") +
  coord_flip()
```

## Questions
- How to check if clustering makes sense? - is 0 values ok?
- What to do about ACTION_TYPE - so many 
- Shots associated with 'winning basketball'
- 


RQs:
- player level: does shot difficulty (distance, shot type, shot zone) predict made shots after accounting for player-level shooting ability?
do physical characteristics (height, weight, position) moderate the distance-accuracy tradeoff?
are taller players more accurate on mid-range vs. close shots compared to guards?

- game level: does shot-making probability change as a function of time remaining (period × minutes/seconds remaining)? do players shoot worse in high-leverage moments (4th quarter, close games, playoffs)? is there a "fatigue effect" — do later periods show lower FG% after controlling for shot type and distance?

- team level: does team quality (TEAM_W_PCT, TEAM_PLUS_MINUS) predict individual shot-making above and beyond player ability?
do players on better teams take higher-quality shots (shorter distance, better zones)?
  
- shot level: shot  do certain ACTION_TYPEs (pullup jumpers vs. catch-and-shoot) have different success rates after controlling for distance?
are shots from "better" zones (corner 3 vs. above the break) more predictive of makes than raw distance alone?


## ICC

scaling data? maybe help w optimization?
```{r}
nba_scale <- nba %>%
  mutate(across(where(is.numeric), ~as.numeric(scale(.))))
```


could look for some resource online on ICC for logistic

```{r}
nba_glmer_base_0 <- glmer(factor(SHOT_MADE_FLAG) ~ 1 + (1|PLAYER_ID), family=binomial, data=nba) 
summary(nba_glmer_base_0)
```


```{r}
nba_glmer_base <- glmer(factor(SHOT_MADE_FLAG) ~ 1 + (1|PLAYER_ID) + (1|GAME_ID), 
                        family = binomial, data = nba, 
                        nAGQ = 0,
                        control = glmerControl(
                          optCtrl = list(maxfun = 1e6),
                          tol = 1e-3
                        ))

summary(nba_glmer_base)
```

#### ICC Latent Threshold Approach:
(https://online.ucpress.edu/collabra/article/10/1/94263/200117/Approaches-for-Quantifying-the-ICC-in-Multilevel)

Player:
```{r}
tau20 <- VarCorr(nba_glmer_base)$PLAYER_ID[1]
sigma2 <- pi^2/3
threshholdICC = tau20 / (tau20 + sigma2)
threshholdICC
```

Game:
```{r}
tau20 <- VarCorr(nba_glmer_base)$GAME_ID[1]
sigma2 <- pi^2/3
threshholdICC = tau20 / (tau20 + sigma2)
threshholdICC
```



## Big model
```{r}
nba_glmer <- glmer(factor(SHOT_MADE_FLAG) ~ SHOT_CATEGORY * PERIOD + SHOT_CATEGORY*GAME_TYPE + SHOT_CATEGORY*SHOT_DISTANCE + 
                     AGE + HEIGHT + WEIGHT + POSITION*SHOT_DISTANCE + TEAM_W + (SHOT_DISTANCE|PLAYER_ID) + (1|GAME_ID), 
                        family = binomial, data = nba, 
                        nAGQ = 0,
                        control = glmerControl(
                          optCtrl = list(maxfun = 1e6),
                          tol = 1e-3
                        ))
summary(nba_glmer)
```






